\center \fontsize{18.8}{20} \selectfont Towards Globally Color-Dual All-Loop Integrands

### 2023-09-18 notes/2loopNLSM.tex
edits

Aside from a few special theories that permit an off-shell Lagrangian description \citepls{}, color-dual integrands are constructed from an ansatz. The four conditions we impose on an ansatz are summarized here and elaborated on below.\footnote{The four ansatz conditions are similar to those of Ref.~\cite{Bern:2015ooa} except that the authors demand that the numerators have the same loop power counting as Feynman gauge Feynman rules.  This is an additional assumption that is most emphatically not made here.}
\begin{enumerate}
\item \textbf{Locality}: Diagram numerators are polynomials in momenta and
  polarizations, \eg the diagrams only have poles given by the
  propagators of the diagrams.
  \item  \textbf{Color-kinematics duality}: The numerators obey the same Jacobi identities as the color factors.
This implicitly requires that the integrand is expressed in terms of purely cubic diagrams.
\item  \textbf{Graph symmetries}: The diagram numerators are only a function of the diagram's topology and labeling.  Furthermore, the diagram numerators are invariant under the automorphisms of the diagrams \emph{including signs that compensate for color-factor sign changes}.
\item  \textbf{Unitarity cuts}: The cuts of the ansatz must reproduce the physical unitarity cuts of the theory.
\end{enumerate}

@@

\paragraph{Locality}

@@

\paragraph{Unitarity}

@@

\subsection{One-loop color-dual integrands}
Our first application of this theory for color-dual construction at loop level is for self-dual Yang-Mills (SDYM). At tree-level, the off-shell cubic vertices can be written in terms of light cone coordinates \cite{}:
\be
X(p,k) = p_u k_w-p_w k_u
\ee
At loop-level, this construction needs to be analytically continued to general dimension in order to apply dimensional regularization at one-loop. This can be done using the cubic semi-YM vertices of the previous section:
\be
\cubic{hgrey0}{fermion2}{}{fermion2}{}{fermion2}{}= \mathcal{X}(k_1,k_2) = (\varepsilon_1\varepsilon_3)(\varepsilon_2 k_3) -  (\varepsilon_2\varepsilon_3)(\varepsilon_1 k_3)
\ee
Plugging in on-shell all-plus helicity states in light-cone gauge will yield precisely the 4D SDYM vertex, up to an unphysical phase. Of course, this has the advantage of permitting a $D$-dimensional construction of the one-loop integrand. Below we provide a four-point example:
\be
\simplebox = g_{\text{YM}}^4 \langle \mathcal{X}(k_1,l_1)\mathcal{X}(k_2,l_2)\mathcal{X}(k_3,l_3)\mathcal{X}(k_4,l_4)\rangle
\ee
where $l_i = l+k_1+k_2+\cdots+ k_i$. In general, the $n$-gon vertex would be
\be
\mathcal{N}^{\text{SDYM}}_{n\text{-gon}}= \nGonVector = g_{\text{YM}}^n \langle \mathcal{X}(k_1,l_1)\mathcal{X}(k_2,l_2)\cdots \mathcal{X}(k_n,l_n) \rangle
\ee
Due to the state-sum of internal loop factors, $\sum {\varepsilon_{(+l)}\varepsilon_{(-l)}}=D$, the integrand above depends explicitly on the spacetime dimension, $D$. By taking a derivative, we can recover the integrand needed for all-plus one-loop amplitudes with an internal scalar:
\be
\ee
Note from James:  I've explicitly checked the nice form of the 1-loop numerators at 4pt, 6pt, and 8pt.
\be
\nGonScalar \equiv \frac{\partial}{\partial D} \mathcal{N}^{\text{SDYM}}_{n\text{-gon}} = (2 g_{\text{YM}})^n (l_1 k_1)(l_2 k_2)\cdots (l_n k_n)
\ee
This integrand numerator with internal scalar loop is precisely what one would obtain from the Feynman rules of YZ-theory, absent the $XZ$ internal vector loop. For more background, we refer the reader to appendix \ref{}. 
\be
\left[\nGonScalar\right]_{\epsilon \rightarrow k} =  f_\pi^{-n} [12][23]\cdots [n1]
\ee
Where we have define the antisymmetric kinematic variable, $[LR] = l_L^2 - l_R^2$
\subsection{Two-loop obstruction}
At two-loop, introducing terms that conspire with four-point contacts is unavoidable. Thus, to approach the problem of two-loop construction, we employ an ansatz based approach. 
\section{Two-loop bootstrap}

@@

\section{Conclusions and Outlook}
\subsection{All-loop No-Go Theorem}
This could be a section about our expectations for all-loop construction requiring non-local color-dual integrands. Lay out counterfactual proof, i,e. define the following statement:
\be
 Q_L=\textit{``the $L$-loop numerator is polynomial and color-dual"}. 
\ee
Then it is clear that if the kinematic numerator is polynomial and color-dual, evaluating it in kinematic regions where the internal legs are on-shell won't disrupt the kinematic algrebraic relations, and it must still be polynomial. Thus we find:
\be
Q_L \Rightarrow Q_{L-1}
\ee
Or equivalently,
\be
\text{not }Q_{L-1}\Rightarrow \text{not }Q_L
\ee
We have found that the two-loop integrand for Yang-Mills cannot be constructed from a color-dual polynomial kinematic numerator. Thus, if a color-dual representation exists at all-loop order for Yang-Mills it must be a rational function of kinematics.

Much of the complexity for $D$-dimensional vector theories at loop-level arises due to the mixing between external polarization and internal loop momenta. One advantage of expanding ansatz construction towards functions of rational kinematics is the prospect of stripping out overall factors of gauge invariant tensors. For example, consider the simple four-point example of a color-dual representation of Yang-Mills theory. 
\be
\mathcal{N}_{(12|34)}^{\text{YM}} = \frac{1}{3}\left[\frac{t_8F^4}{s_{23}} -  \frac{t_8F^4}{s_{13}}\right]
\ee
\be
\mathcal{N}_{(12|34)}^{\text{YM}} + \mathcal{N}_{(13|42)}^{\text{YM}} +\mathcal{N}_{(14|23)}^{\text{YM}} = 0
\ee
\be
\text{Res}\left[\frac{c_s \mathcal{N}_{s}^{\text{YM}} }{s}\right]^{s=0}= \frac{2}{3} \frac{t_8F^4}{t} \qquad 
\text{Res}\left[\frac{\mathcal{N}_{t}^{\text{YM}} }{t}\right]^{s=0}= \frac{1}{3} \frac{t_8F^4}{t}
\ee
\be
A^{\text{YM}}_{(1234)} = \frac{t_8F^4}{s_{12}s_{23}}
\ee
\subsection{Future Directions}\label{sec:Discussion}
\paragraph{Acknowledgments} The authors would like to thank ... for insightful conversations, related collaboration, and encouragement throughout the completion of this work. This work was supported by the DOE under contract DE-SC0015910 and by the Alfred P. Sloan Foundation. N.H.P. additionally acknowledges the Northwestern University Amplitudes and Insight group, the Department of Physics and Astronomy, and Weinberg College for their generous support.
Feynman diagrams were typeset using TikZ-Feynman \cite{Ellis:2016jkw}.

>>>>

Aside from a few special theories that permit an off-shell Lagrangian description \citepls{}, color-dual integrands are constructed from an ansatz. The four conditions we impose on an ansatz are summarized here and elaborated on below.\footnote{The four ansatz conditions are similar to those of Ref.~\cite{Bern:2015ooa} except that the authors demand that the numerators have the same loop power counting as Feynman gauge Feynman rules.  This is an additional assumption that is most emphatically not made here.}
\begin{enumerate}
\item \textbf{Off-shell Locality}: Diagram numerators are polynomials in momenta and
  polarizations, \eg the diagrams only have poles given by the
  propagators of the diagrams.
  \item  \textbf{Color-kinematics duality}: The numerators obey the same Jacobi identities as the color factors.
This implicitly requires that the integrand is expressed in terms of purely cubic diagrams.
\item  \textbf{Graph symmetries}: The diagram numerators are only a function of the diagram's topology and labeling.  Furthermore, the diagram numerators are invariant under the automorphisms of the diagrams \emph{including signs that compensate for color-factor sign changes}.
\item  \textbf{On-shell Unitarity }: The cuts of the ansatz must reproduce the physical unitarity cuts of the theory when internal momenta are taken on-shell.
\end{enumerate}

@@

\paragraph{Off-shell Locality}

@@

\paragraph{On-shell Unitarity}

@@

\subsection{One-loop color-dual integrands}
Our first application of this theory for color-dual construction at loop level is for self-dual Yang-Mills (SDYM). At tree-level, the off-shell cubic vertices can be written in terms of light cone coordinates \cite{}:
\be
X(p,k) = p_u k_w-p_w k_u
\ee
At loop-level, this construction needs to be analytically continued to general dimension in order to apply dimensional regularization at one-loop. This can be done using the cubic semi-YM vertices of the previous section:
\be
\cubic{hgrey0}{fermion2}{}{fermion2}{}{fermion2}{}= \mathcal{X}(k_1,k_2) = (\varepsilon_1\varepsilon_3)(\varepsilon_2 k_3) -  (\varepsilon_2\varepsilon_3)(\varepsilon_1 k_3)
\ee
Plugging in on-shell all-plus helicity states in light-cone gauge will yield precisely the 4D SDYM vertex, up to an unphysical phase. Of course, this has the advantage of permitting a $D$-dimensional construction of the one-loop integrand. Below we provide a four-point example:
\be
\simplebox = g_{\text{YM}}^4 \langle \mathcal{X}(k_1,l_1)\mathcal{X}(k_2,l_2)\mathcal{X}(k_3,l_3)\mathcal{X}(k_4,l_4)\rangle
\ee
where $l_i = l+k_1+k_2+\cdots+ k_i$. In general, the $n$-gon vertex would be
\be
\mathcal{N}^{\text{SDYM}}_{n\text{-gon}}= \nGonVector = g_{\text{YM}}^n \langle \mathcal{X}(k_1,l_1)\mathcal{X}(k_2,l_2)\cdots \mathcal{X}(k_n,l_n) \rangle
\ee
Due to the state-sum of internal loop factors, $\sum {\varepsilon_{(+l)}\varepsilon_{(-l)}}=D$, the integrand above depends explicitly on the spacetime dimension, $D$. By taking a derivative, we can recover the integrand needed for all-plus one-loop amplitudes with an internal scalar:
\be
\frac{\partial}{\partial D} \mathcal{N}^{\text{SDYM}}_{n\text{-gon}}=\nGonScalar  = (2 g_{\text{YM}})^n (l_1 k_1)(l_2 k_2)\cdots (l_n k_n)
\ee
This integrand numerator with internal scalar loop is precisely what one would obtain from the Feynman rules of YZ-theory, absent the $XZ$ internal vector loop. For more background, we refer the reader to appendix \ref{}. 
\be
 \mathcal{N}^{\text{NLSM}}_{n\text{-gon}}=\left[\frac{\partial}{\partial D} \mathcal{N}^{\text{SDYM}}_{n\text{-gon}}\right]^{\epsilon \rightarrow k}=\nGonScalar =  f_\pi^{-n} [12][23]\cdots [n1]
\ee
Where we have define the antisymmetric kinematic variable, $[LR] = l_L^2 - l_R^2$
\subsection{Two-loop obstruction}
At two-loop, introducing terms that conspire with four-point contacts is unavoidable (add diagram). Thus, to approach the problem of two-loop construction, we employ an ansatz based approach. 
\section{Two-loop bootstrap}

@@

\section{Conclusions and Outlook}
In summary, we have found that the two-loop integrand for Yang-Mills cannot be constructed from a color-dual numerator that is a polynomial in the kinematic variables. Thus, if a color-dual representation exists at all-loop order for Yang-Mills, it must be a rational function of kinematics.

This seems to be a rather unnatural constraint on color-dual theories from the perspective of point-like quantum field theories. After all, introducing operators that produce rational functions of kinematics must be non-local. Thus, in order to identify valid color-dual representation of Yang-Mills beyond tree-level, we must relax the constraint of \textit{off-shell locality}, and realize locality only in the limit of \textit{on-shell} kinematics. Below we provide an overview of what one might expect for such a construction, and provide some simple examples at tree-level that manifest this structure. 
\subsection{Non-local construction of scattering amplitudes}
 For example, consider the simple four-point example of a color-dual representation of Yang-Mills theory. 
\mathcal{N}_{(12|34)}^{\text{YM}} = \frac{t_8F^4}{3} \frac{s_{13}-s_{12}}{s_{23}{s_{13}}}
By construction, this functionally symmetric numerator is anti-symmetric and obeys the Jacobi identity:
\mathcal{N}_{(12|34)}^{\text{YM}} + \mathcal{N}_{(13|42)}^{\text{YM}} +\mathcal{N}_{(14|23)}^{\text{YM}} = 0
However, it must also factorize to kinematics that are consistent with the local Feynman rules of \ref{}. Evaluating the residues on $s=s_{12}$, $t=s_{23}$ and $u=s_{13}$ poles, we find:
\text{Res}\left[\frac{ \mathcal{N}_{s}^{\text{YM}} }{s}\right]^{s=0}&= \frac{2}{3}A_{(12|l)}^{\text{YM}}A_{(-l|34)}^{\text{YM}} 
\text{Res}\left[\frac{\mathcal{N}_{t}^{\text{YM}} }{t}\right]^{s=0}&= -\frac{1}{3}A_{(12|l)}^{\text{YM}}A_{(-l|34)}^{\text{YM}} 
\text{Res}\left[\frac{\mathcal{N}_{u}^{\text{YM}} }{u}\right]^{s=0}&= -\frac{1}{3} A_{(12|l)}^{\text{YM}}A_{(-l|34)}^{\text{YM}} 
Plugging these into a cubic graph representation of Yang-Mills \ref{}, we find:
\text{Res}\left[\mathcal{A}_4^{\text{YM}}\right]^{s=0} = \frac{1}{3} A_{(12|l)}^{\text{YM}}A_{(-l|34)}^{\text{YM}} (2c_s-c_t-c_u) =  \mathcal{A}_{(12|l)}^{\text{YM}}\mathcal{A}_{(-l|34)}^{\text{YM}} 
where we have applied the color structure Jacobi identity, $c_s+c_t+c_u=0$. While we can absolutely apply a generalized gauge trasformation \cite{BCJ} to restore field theoretic locality to the cubic numerators, we can see clearly that this is an aesthetic choice. Moreover, by abandoning traditional notions of off-shell locality, we can massage the color-dual numerators into a form that is manifestly gauge invariant. As an organizational principle, pulling out an overall gauge-invariant factors comes with the added advantage of simplifying a loop-level construction of color-dual numerators for Yang-Mills. 
\subsection{Non-local construction of color-dual integrands}
Much of the complexity for $D$-dimensional vector theories at loop-level arises due to the mixing between external polarization and internal loop momenta. As noted, one advantage of expanding ansatz construction towards functions of rational kinematics is the prospect of stripping out overall factors of gauge invariant tensors. Rather than building an ansatz from the irreducible scalar products of \ref{}, 
\mathcal{N}^{\text{YM}}_{m,L} = \sum_i \mathcal{O}_i \mathcal{R}^{(i)}_{m,L}
where $\mathcal{O}_i$ are the basis tensors of \cite{}, and $ \mathcal{R}_i$ is a rational function of momenta of the form;
\mathcal{R}_{m,L} = \frac{P_{b+a}(kk,kl,ll)}{Q_a(kk,kl,ll)}, \qquad b=m+2(l-1)
where $P_{a}$ and $Q_a$ are degree-$a$ polynomials exclusively of $D$-dimensional scalar kinematics. 
\subsection{Future Directions}\label{sec:Discussion}
\paragraph{Acknowledgments} The authors would like to thank ... for insightful conversations, related collaboration, and encouragement throughout the completion of this work. This work was supported by the DOE under contract DE-SC0015910 and by the Alfred P. Sloan Foundation. N.H.P. additionally acknowledges the Northwestern University Amplitudes and Insight group, the Department of Physics and Astronomy, and Weinberg College for their generous support.
Feynman diagrams were typeset using TikZ-Feynman \cite{Ellis:2016jkw}.
